---
title: "Case Study 2"
author: "Lauren Nelson | Jordan Eaddy"
date: "12/11/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DataExplorer) #Simplified EDA 
library(readr)
library(dplyr) #filtering
library(ggplot2)
library(tidyverse)
library(caret)
library(class)
library(e1071)#Naive Bayes
library(tidyr)
library(stringr)
library(plotly)
library(psych)
library(lessR)
library(GGally)
library(ggcorrplot)
library (ggthemes)
library(inspectdf)
library(scales)
library(readxl)
library(Metrics)
library(DT)
```

##Description
DDSAnalytics is an analytics company that specializes in talent management solutions for Fortune 1000 companies. Talent management is defined as the iterative process of developing and retaining employees. It may include:

-workforce planning
-employee training programs 
-identifying high-potential employees and 
-reducing/preventing voluntary employee turnover (attrition) 

To gain a competitive edge over its competition, DDSAnalytics is planning to leverage data science for talent management. The executive leadership has identified predicting employee turnover as its first application of data science for talent management. Before the business green lights the project, they have tasked your data science team to conduct an analysis of existing employee data.


```{r}
#read in data
CSD <- read.csv("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%2014%20and%2015%20Case%20Study%202/CaseStudy2-data.csv", header = T)

NoAttr <- read.csv("/Users/lnelson/Box/MSDS/MSDS_6306_Doing-Data-Science-Master/Unit 14 and 15 Case Study 2/CaseStudy2CompSet No Attrition.csv")

Salary <- read_excel("/Users/lnelson/Box/MSDS/MSDS_6306_Doing-Data-Science-Master/Unit 14 and 15 Case Study 2/CaseStudy2CompSet No Salary.xlsx")


glimpse(CSD) # Show Data Structure
names <- c(3,4,6,9,13,17,19,23,24) #Categorical Variables
head(names)
CSD[,names] <-lapply(CSD[,names] , factor) #convert chr to factor
as.numeric(as.character(names))
glimpse(CSD) # Show Data Structure

CSD$Attrition<-ifelse(CSD$Attrition=="Yes",1,0) # Converted Attrition to a logical operation. 1 = Yes, 0 = No
CSD$Attrition<-as.numeric(as.character(CSD$Attrition))
CSD$OverTime<-ifelse(CSD$OverTime=="Yes",1,0) # Converted Overtime to a logical operation, 1 = Yes, 0 = NO
CSD$OverTime<-as.numeric(as.character(CSD$OverTime))


sapply(CSD, class)

#data prep and cleaning
#check CSD for NAs
sapply(CSD,function(x) sum(is.na(x)))
# drop ID, EmployeeCount, EmployeeNumber, Over18, StandardHours -- no value for EDA
CSD <- subset(CSD,select=c(2,3,4,5,6,7,8,9,12,13,14,15,16,17,18,19,20,21,22,24,25,26,27,29,30,31,32,33,34,35,36))
sapply(CSD, class)
#Distance from Home
#EnvironmentSatisfaction
#PerformanceRating
```

#Exploratory Data Analysis

```{r exploratory data analysis}
#structure of the data
str(CSD)
#dimensions
dim(CSD)
#high-level view of continuous variables
plot_histogram(CSD, binary_as_factor = TRUE, geom_histogram_args = list(bins = 30L), scale_x = "continuous")
plot_density(CSD)
#Categorical Variables-Barplots
plot_bar(CSD)
```

#Interesting Insights into the Data
```{r}
#Data
cordata = CSD[,c(1,2,4,6,7,9,11,12,13,15,17,18,19,21,22,23,24,25,26,27,28,29,30,31)] #Numerical Variables
corr <- round(cor(cordata), 1) #Create Correlation Matrix
#corr #Correlation Matrix

#Job Level and Monthly Income = 1
#Total Working Years and Job level = .8
#Total Working Years and Monthly Income = .8
#Percent Salary Hike and Performance Rating = .8
#Age and Total Working Years = .7

#Indicates a linear relationship between Total Working Years and JobLevel
plot(CSD$TotalWorkingYears,CSD$JobLevel, main="Correlation between Total Working Years & Job Level", xlab="Total Working Years", ylab="Job Level")

cor(CSD$TotalWorkingYears, CSD$JobLevel)

cor.test(CSD$TotalWorkingYears, CSD$JobLevel)

#Indicates a linear relationship between Total Working Years and Monthly Income
plot(CSD$TotalWorkingYears,CSD$MonthlyIncome, main="Correlation between Total Working Years & Monthly Income", xlab="Total Working Years", ylab="Monthly Income")

cor(CSD$TotalWorkingYears,CSD$MonthlyIncome)

cor.test(CSD$TotalWorkingYears,CSD$MonthlyIncome)

#Indicates a linear relationship between Percent Salary Hike and Performance Rating
plot(CSD$PercentSalaryHike,CSD$PerformanceRating, main="Correlation between Percent Salary Hike & Performance Rating", xlab="Percent Salary Hike", ylab="Performance Rating")

cor(CSD$PercentSalaryHike,CSD$PerformanceRating)

cor.test(CSD$PercentSalaryHike,CSD$PerformanceRating)

#Relationship between Categorical Variables
#Frequency Table
# 2 - way table


#Attrition by OverTime
BarChart(Attrition, data = CSD, by=OverTime, stack100=TRUE, main = "Attrition by OverTime", outer = TRUE, fill=c("black", "gold3"))

#2 Marital_Status-Single
two_wayMA = table(CSD$MaritalStatus, CSD$Attrition)

two_wayMA


prop.table(two_wayMA) # cell percentages

#More Attrition among Single Marital Status
BarChart(Attrition, data = CSD, by=MaritalStatus, stack100=TRUE, main = "Attrition by Marital Status", outer = TRUE, fill=c("black", "gold3", "turquoise3"))

#3 Job Role Sales Representative
two_wayJS = table(CSD$JobRole, CSD$Attrition)

two_wayJS


prop.table(two_wayJS) # cell percentages

#Attrition by Job Role
BarChart(Attrition, data = CSD, by=JobRole, stack100=TRUE, main = "Attrition by Job Role", outer = TRUE, )

BarChart(Attrition, data = CSD)
BarChart(JobRole, data = CSD)
#Job Role by Gender
BarChart(JobRole, by=Gender, stack100=TRUE, main = "Job Role by Gender", data = CSD, fill = c("pink3","blue3"))
#Distribution of Marital Status
BarChart(MaritalStatus, data = CSD, legend_labels = TRUE)
#Marital Status Distribution by Gender
BarChart(MaritalStatus, by=Gender, stack100=TRUE, main = "Marital Status by 
Gender", data = CSD, fill = c("pink3","blue3"))
BarChart(OverTime, data = CSD)
BarChart(EducationField, data = CSD, by1=Attrition)
#BarChart(Attrition, data = CSD, by1=BusinessTravel)
BarChart(Attrition, data = CSD, by=BusinessTravel, stack100=TRUE, main = "Attrition by Business Travel", outer = TRUE)
#More Attrition among Single Marital Status
BarChart(Attrition, data = CSD, by=MaritalStatus, stack100=TRUE, main = "Attrition by Marital Status", outer = TRUE, fill=c("turquoise", "blue3", "forestgreen"))
#Attrition by Gender
BarChart(Attrition, data = CSD, by=Gender, stack100=TRUE, main = "Attrition by Gender", outer = TRUE, fill=c("pink3", "blue3"))


#Attrition by Job Satisfaction - lower job satisfaction = increased attrition
BarChart(JobSatisfaction, data = CSD, by=Attrition, stack100=TRUE, main = "Attrition by Job Satisfaction", outer = TRUE)
```


```{r FINAL Insight: Attrition}
CSD %>% 
  group_by(Attrition, Gender) %>% 
  summarise(Attrition_Percent= 100 * n()/ nrow(CSD))
# ** Only 16 % of our data actually left their position for one reason or another**
# Note to self: This may cause problems in the modeling phase. Consider sampling ideas. 

#JobRole Trends

CSD %>%  # Viz
  ggplot(aes(JobRole, fill = JobRole))+
  geom_bar(aes(y=(..count..)/sum(..count..)))+
  theme(legend.position = "None")+
  scale_y_continuous(labels= percent)+
    labs(title= "Job Role Percentages", y = "Percentage", x = "Job Role",tag = "Fig. 1" )+
  theme_igray()

# Income by Job Title Distributions 
CSD %>% 
  ggplot(aes(reorder(x=JobRole,-MonthlyIncome),y=MonthlyIncome, fill= JobRole))+
  geom_boxplot()+
  scale_y_continuous(labels= dollar_format())+
  labs(title= "Income by Role", y = "Income", x = "Job Title", tag = "Fig. 3")+
  theme_economist_white()+
  theme(legend.position = "none",axis.title.y = element_text(vjust= 4) )+
  guides(x = guide_axis(n.dodge= 2))

# 16 Percent left thier role
CSD %>% 
  group_by(Attrition) %>% 
  summarise(Attrition_Percent= 100 * n()/ nrow(CSD))

# Attrition by Job Role
CSD %>% 
  ggplot(aes(JobRole, y=(..count..)/sum(..count..), fill=Attrition))+
  geom_bar()+
  theme_economist_white()+
  theme(axis.title.x = element_text(vjust=-4))+
  labs(title= "Attrition By Job Role", x = "Job Role", y= "", tag= "Fig. 2")+
  scale_y_continuous(labels = percent)

# Sales Reps have high attrition, despite having a Job Satisfaction rating in the upper percentile of our data. Monthly Income
CSD%>%
  ggplot(aes(Gender))+
  geom_histogram(stat="count")+
  facet_grid(~JobSatisfaction)+
  labs(title= "Job Satisfaction by Gender", y="", tag= "Fig. 3")+
  theme_economist_white()+
  theme(axis.text.y = element_blank())

# Women seem to be not satisfied with their work situation
CSD %>% 
  ggplot(aes(Gender, MonthlyIncome))+
  geom_bar(stat="identity")+
  facet_wrap(~JobRole)+
  scale_y_continuous(labels=dollar_format())+
  labs(title = "Income by Gender & Job Role", tag = "Fig.4",y="Monthly Income")+
  theme_economist_white()+
  theme(axis.title.y = element_text(vjust = 2))
 
# Women are paid less than men
CSD %>% 
  ggplot(aes(Gender,Education, fill = factor(Education)))+
  geom_bar(stat="identity", position = "dodge")+
  facet_wrap(~JobRole)+
  theme_economist_white()+
  theme(axis.title.y = element_text(vjust = 2))+
  labs(title = "Education Levels by Job Role", y="Education Level", tag="Fig.3")
```

#Identify Top 3 factors that lead to attrition

```{r}
#### We will now start looking for the correlations in the data between variables. 
#### correlations between variables
#continuous
plot_correlation(CSD, type = "continuous", geom_text_args = list(),theme_config = list(legend.position = "right", axis.text.x = element_text(angle =
    90)))
#discrete
plot_correlation(CSD, type = "discrete", geom_text_args = list(),theme_config = list(legend.position = "right", axis.text.x = element_text(angle =
    90)))
#1 OverTime
#2 Marital_Status-Single
#3 Job Role Sales Representative
```


\#Identify Job Role specific trends 

```{r}
#Attrition by Job Satisfaction - lower job satisfaction = increased attrition
BarChart(JobRole, data = CSD, by1=JobSatisfaction)
```


KNN - Overview of Process

1.  Decide on *similarity* or *distance* metric
2.  Split the original labeled dataset into training and test data
3.  Pick an evaluation metric
4.  Run KNN a few times, changing *k* and checking the evaluation measure
5.  Optimize *k* by picking the one with the best evaluation measure
6.  Once *k* is chosen, use the same training set and now create a new test set with No Attrition "CaseStudy2CompSet No Attrition.csv" \| then with No Salary "CaseStudy2CompSet No Salary.csv"

#Output should be "Case2PredictionsNelson-EaddyAttrition.csv -- add to github repo


```{r echo=TRUE}
#install.packages("caTools")
library(caTools)


#Classify Attrition into two classes: 
#1. True
#2. False

attrition.subset <- CSD[c(6,20,22)]
#CSD[,attrition.subset$Attrition] <-lapply(CSD[,attrition.subset$Attrition] , factor) #convert chr to factor
#CSD$Attrition<-ifelse(CSD$Attrition=="Yes",1,0)
#CSD$Attrition<-as.numeric(as.character(CSD$Attrition))
#Normalization
normalize <- function(x) {
return ((x - min(x)) / (max(x) - min(x))) }
attrition.subset.n <- as.data.frame(lapply(attrition.subset[,1:3], normalize))

#Data Splicing
set.seed(5)
attr.d <- sample(1:nrow(attrition.subset.n),size=nrow(attrition.subset.n)*0.7,replace = FALSE) #random selection of 70% data.
 
train.attrition <- attrition.subset[attr.d,] # 70% training data
test.attrition <- attrition.subset[-attr.d,] # remaining 30% test data

#Creating seperate dataframe for 'Attrition' feature which is our target.
train.attrition_labels <- attrition.subset[attr.d,2]
test.attrition_labels <-attrition.subset[-attr.d,2]


	#Find the number of observation
NROW(train.attrition_labels) 
sqrt(609) #25
knn.25 <- knn(train=train.attrition, test=test.attrition, cl=train.attrition_labels, k=25)
knn.11 <- knn(train=train.attrition, test=test.attrition, cl=train.attrition_labels, k=11)

#Model Evaluation
#Calculate the proportion of correct classification for k = 25, 1
ACC.25 <- 100 * sum(test.attrition_labels == knn.25)/NROW(test.attrition_labels)
ACC.25
ACC.11 <- 100 * sum(test.attrition_labels == knn.11)/NROW(test.attrition_labels)
ACC.11

confusionMatrix(table(knn.25 ,test.attrition_labels))
confusionMatrix(table(knn.11 ,test.attrition_labels))

i=1
k.optm=1
for (i in 1:30) {
 knn.mod <- knn(train=train.attrition,test=test.attrition, cl=train.attrition_labels, k=i)
 k.optm[i] <- 100 * sum(test.attrition_labels == knn.mod)/NROW(test.attrition_labels)
 k=i
 cat(k,'=',k.optm[i],'')
}

```

```{r Income Model}
setwd("/Users/lnelson/Box/MSDS/MSDS_6306_Doing-Data-Science-Master/Unit 14 and 15 Case Study 2/")
RMSE<-function(error) {sqrt(mean(error^2))} #Function to get RMSE programatically 
Incomesplit<- .80
SM<-CSD%>% 
  select(JobLevel,TotalWorkingYears, MonthlyIncome, Age, JobRole)
# Model Creation
Trainindicieslm= sample(1:dim(SM)[1], round(Incomesplit * dim(SM)[1]))
LRMtr<- SM[Trainindicieslm,]
LRMte<- SM[-Trainindicieslm,]
fit=lm(MonthlyIncome~., data=LRMtr)
# Model Creation
# Model Test RMSE < 3000
prediction<-predict.lm(fit, newdata = Salary)
rmse(LRMte$MonthlyIncome, predict(fit, LRMte)) #Sanity Check
# 935.93
summary(prediction)
predictiondf<-as.data.frame(prediction)
Salary$Salary<-predictiondf[,1]
CSPRED = Salary %>% 
  select("ID","Salary") %>% 
  arrange(ID)
write.csv(CSPRED, file = "Case2Predictions_Nelson_Eaddy Salary.csv")
```


```{r}
#NaiveBayes model  

NBCSD = CSD %>% filter(!is.na(DistanceFromHome) & !is.na(OverTime) & !is.na(PerformanceRating) & !is.na(Attrition))
set.seed(9)
trainIndices1 = sample(seq(1:length(NBCSD$Attrition)),round(.7*length(NBCSD$Attrition)))
train1 = NBCSD[trainIndices1,]
test1 = NBCSD[-trainIndices1,]

head(train1)
head(test1)

model1 = naiveBayes(train1[,c(6,20,22)],(train1$Attrition),laplace = 1)

table(predict(model1,test1[,c(6,20,22)]),(test1$Attrition))

CM5 = confusionMatrix(table(predict(model1,test1[,c(6,20,22)]),(test1$Attrition)))

CM5
```

Create a full EDA report

```{r full EDA report, echo=FALSE}
#Create a full EDA report
#create_report(data=CSD, output_file = 'Case Study 2 EDA.html', y='Attrition')
```

